### 2023-01-02 paper/paper.tex
Update paper.tex

\pdfoutput=1
\documentclass[11pt]{article}
\usepackage{acl}
% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}

\title{AICS project working title}

\author{Dominik Künkele \\
  Master in Language Technology \\
  University of Gothenburg \\
  \texttt{guskunkdo@student.gu.se} \\\And
  Maria Irena Szawerna \\
  Master in Language Technology \\
  University of Gothenburg \\
  \texttt{gusszawma@student.gu.se} \\}

\begin{document}
\maketitle
\begin{abstract}
This document is a supplement to the general instructions for *ACL authors. It contains instructions for using the \LaTeX{} style files for ACL conferences. 
The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
These instructions should be used both for papers submitted for review and for final versions of accepted papers.
\end{abstract}

\section{Introduction}

These instructions are for authors submitting papers to *ACL conferences using \LaTeX. They are not self-contained. All authors must follow the general instructions for *ACL proceedings,\footnote{\url{http://acl-org.github.io/ACLPUB/formatting.html}} and this document contains additional instructions for the \LaTeX{} style files.

The templates include the \LaTeX{} source of this document (\texttt{acl.tex}),
the \LaTeX{} style file used to format it (\texttt{acl.sty}),
an ACL bibliography style (\texttt{acl\_natbib.bst}),
an example bibliography (\texttt{custom.bib}),
and the bibliography for the ACL Anthology (\texttt{anthology.bib}).

\section{Engines}

To produce a PDF file, pdf\LaTeX{} is strongly recommended (over original \LaTeX{} plus dvips+ps2pdf or dvipdf). Xe\LaTeX{} also produces PDF files, and is especially suitable for text in non-Latin scripts.

\section{Preamble}

The first line of the file must be
\begin{quote}
\begin{verbatim}
\documentclass[11pt]{article}
\end{verbatim}
\end{quote}

To load the style file in the review version:
\begin{quote}
\begin{verbatim}
\usepackage[review]{acl}
\end{verbatim}
\end{quote}
For the final version, omit the \verb|review| option:
\begin{quote}
\begin{verbatim}
\usepackage{acl}
\end{verbatim}
\end{quote}

To use Times Roman, put the following in the preamble:
\begin{quote}
\begin{verbatim}
\usepackage{times}
\end{verbatim}
\end{quote}
(Alternatives like txfonts or newtx are also acceptable.)

Please see the \LaTeX{} source of this document for comments on other packages that may be useful.

Set the title and author using \verb|\title| and \verb|\author|. Within the author list, format multiple authors using \verb|\and| and \verb|\And| and \verb|\AND|; please see the \LaTeX{} source for examples.

By default, the box containing the title and author names is set to the minimum of 5 cm. If you need more space, include the following in the preamble:
\begin{quote}
\begin{verbatim}
\setlength\titlebox{<dim>}
\end{verbatim}
\end{quote}
where \verb|<dim>| is replaced with a length. Do not set this length smaller than 5 cm.

\section{Document Body}

\subsection{Footnotes}

Footnotes are inserted with the \verb|\footnote| command.\footnote{This is a footnote.}

\subsection{Tables and figures}

See Table~\ref{tab:accents} for an example of a table and its caption.
\textbf{Do not override the default caption sizes.}

\begin{table}
\centering
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\"a}| & {\"a} \\
\verb|{\^e}| & {\^e} \\
\verb|{\`i}| & {\`i} \\ 
\verb|{\.I}| & {\.I} \\ 
\verb|{\o}| & {\o} \\
\verb|{\'u}| & {\'u}  \\ 
\verb|{\aa}| & {\aa}  \\\hline
\end{tabular}
\begin{tabular}{lc}
\hline
\textbf{Command} & \textbf{Output}\\
\hline
\verb|{\c c}| & {\c c} \\ 
\verb|{\u g}| & {\u g} \\ 
\verb|{\l}| & {\l} \\ 
\verb|{\~n}| & {\~n} \\ 
\verb|{\H o}| & {\H o} \\ 
\verb|{\v r}| & {\v r} \\ 
\verb|{\ss}| & {\ss} \\
\hline
\end{tabular}
\caption{Example commands for accented characters, to be used in, \emph{e.g.}, Bib\TeX{} entries.}
\label{tab:accents}
\end{table}

\subsection{Hyperlinks}

Users of older versions of \LaTeX{} may encounter the following error during compilation: 
\begin{quote}
\tt\verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.
\end{quote}
This happens when pdf\LaTeX{} is used and a citation splits across a page boundary. The best way to fix this is to upgrade \LaTeX{} to 2018-12-01 or later.

\subsection{Citations}

\begin{table*}
\centering
\begin{tabular}{lll}
\hline
\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
\hline
\citep{Gusfield:97} & \verb|\citep| & \verb|\cite| \\
\citealp{Gusfield:97} & \verb|\citealp| & no equivalent \\
\citet{Gusfield:97} & \verb|\citet| & \verb|\newcite| \\
\citeyearpar{Gusfield:97} & \verb|\citeyearpar| & \verb|\shortcite| \\
\hline
\end{tabular}
\caption{\label{citation-guide}
Citation commands supported by the style file.
The style is based on the natbib package and supports all natbib citation commands.
It also supports commands defined in previous ACL style files for compatibility.
}
\end{table*}

Table~\ref{citation-guide} shows the syntax supported by the style files.
We encourage you to use the natbib styles.
You can use the command \verb|\citet| (cite in text) to get ``author (year)'' citations, like this citation to a paper by \citet{Gusfield:97}.
You can use the command \verb|\citep| (cite in parentheses) to get ``(author, year)'' citations \citep{Gusfield:97}.
You can use the command \verb|\citealp| (alternative cite without parentheses) to get ``author, year'' citations, which is useful for using citations within parentheses (e.g. \citealp{Gusfield:97}).

\subsection{References}

\nocite{Ando2005,andrew2007scalable,rasooli-tetrault-2015}

The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
If your own bib file is named \texttt{custom.bib}, then placing the following before any appendices in your \LaTeX{} file will generate the references section for you:
\begin{quote}
\begin{verbatim}
\bibliography{custom}
\end{verbatim}
\end{quote}

You can obtain the complete ACL Anthology as a Bib\TeX{} file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
To include both the Anthology and your own .bib file, use the following instead of the above.
\begin{quote}
\begin{verbatim}
\bibliography{anthology,custom}
\end{verbatim}
\end{quote}

Please see Section~\ref{sec:bibtex} for information on preparing Bib\TeX{} files.

\subsection{Appendices}

Use \verb|\appendix| before any appendix section to switch the section numbering over to letters. See Appendix~\ref{sec:appendix} for an example.

\section{Bib\TeX{} Files}
\label{sec:bibtex}

Unicode cannot be used in Bib\TeX{} entries, and some ways of typing special characters can disrupt Bib\TeX's alphabetization. The recommended way of typing special characters is shown in Table~\ref{tab:accents}.

Please ensure that Bib\TeX{} records contain DOIs or URLs when possible, and for all the ACL materials that you reference.
Use the \verb|doi| field for DOIs and the \verb|url| field for URLs.
If a Bib\TeX{} entry has a URL or DOI field, the paper title in the references section will appear as a hyperlink to the paper, using the hyperref \LaTeX{} package.

\section*{Acknowledgements}

This document has been adapted
by Steven Bethard, Ryan Cotterell and Rui Yan
from the instructions for earlier ACL and NAACL proceedings, including those for 
ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
NAACL 2019 by Stephanie Lukin and Alla Roskovskaya, 
ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu, 
NAACL 2018 by Margaret Mitchell and Stephanie Lukin,
Bib\TeX{} suggestions for (NA)ACL 2017/2018 from Jason Eisner,
ACL 2017 by Dan Gildea and Min-Yen Kan, 
NAACL 2017 by Margaret Mitchell, 
ACL 2012 by Maggie Li and Michael White, 
ACL 2010 by Jing-Shin Chang and Philipp Koehn, 
ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui, 
ACL 2005 by Hwee Tou Ng and Kemal Oflazer, 
ACL 2002 by Eugene Charniak and Dekang Lin, 
and earlier ACL and EACL formats written by several people, including
John Chen, Henry S. Thompson and Donald Walker.
Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.

% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}

\appendix

\section{Example Appendix}
\label{sec:appendix}

This is an appendix.

\end{document}

### 2023-01-02 code/dataset.py
adapted documentation

@@ -24,7 +24,8 @@ END_TOKEN = '<end>'
 PADDING_TOKEN = '<pad>'
 
 class Datasets(Enum):
-    '''???
+    '''By Dominik
+    An enum that lists all possible datasets and can be used instead of magic strings.
     '''
     CLEF = 'imageCLEF'
     FLICKR = 'flickr'
@@ -35,7 +36,7 @@ class Datasets(Enum):
 @dataclass(slots=True, kw_only=True)
 class Sample:
     '''By Dominik.
-    A class which allows for the creation of picture-caption sample objects. 
+    A class which allows for the creation of picture-caption sample objects.
 
     Attributes:
         image_id (str): The ID of the image (file name without the .jpg extension).
@@ -57,11 +58,19 @@ class Sample:
 
 
 class CaptionLoader(ABC):
-    '''???
+    '''By Dominik
+    An abstract CaptionLoader. All of its implementations need to implement the load_captions method
     '''
     @abstractmethod
     def load_captions(self, concat_captions: bool) -> list[Sample]:
-        pass
+        '''An abstract method that allows for loading in captions.
+
+        Args:
+            concat_captions (bool): Defines whether or not captions should be concatenated, in case they consist of more than one distinct phrase.
+
+        Returns:
+            A list of partially filled-out Sample objects.
+        '''
 
 class FlickrCaptionLoader(CaptionLoader):
     '''By Dominik.
@@ -108,7 +117,7 @@ class FlickrCaptionLoader(CaptionLoader):
 
 class CLEFCaptionLoader(CaptionLoader):
     '''By Dominik.
-    A class that allows for loading in the captions and names of associated images from the Flickr dataset.
+    A class that allows for loading in the captions and names of associated images from the imageCLEF dataset.
 
     Attributes:
         annotation_directory (str): The directory in which the file containing the annotations is stored.
@@ -125,7 +134,7 @@ class CLEFCaptionLoader(CaptionLoader):
         '''A method that allows for loading in captions.
 
         Args:
-            concat_captions (bool): Defines whether or not captions should be concatenated, in case they consist of more than one distinct phrase. 
+            concat_captions (bool): Defines whether or not captions should be concatenated, in case they consist of more than one distinct phrase.
             
         Returns:
             A list of partially filled-out Sample objects.
@@ -242,8 +251,10 @@ class ImageDataset(Dataset):
             tokenized_caption = [word.lower()
                                  for word in nltk.word_tokenize(sample.caption)
                                  if word not in punctuations]
-
[…]

### 2023-01-02 code/main.py

@@ -8,6 +8,7 @@ from dataset import (CLEFCaptionLoader, Datasets, FlickrCaptionLoader,
                      ImageDataset)
 from filters import POSRelationFilter, RuleBasedRelationFilter
 from torch.backends import cudnn
+from torch.utils.data import Dataset
 from train import train
 
 device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
@@ -29,7 +30,7 @@ def load_wordmap(word_map_path: str):
         word_map = json.load(j)
     return word_map
 
-def load_data(dataset: str, data_root_dir: str, word_map: dict, number_of_images: int, unknown_filter: float):
+def load_data(dataset: Datasets, data_root_dir: str, word_map: dict, number_of_images: int, unknown_filter: float):
     '''By Dominik.
     A function that allows for the loading in of the data from a specific dataset to be later used in the fine-tuning.
     
@@ -70,7 +71,7 @@ def load_data(dataset: str, data_root_dir: str, word_map: dict, number_of_images
     else:
         raise ValueError('dataset must be one of the following: clef, flickr')
 
-def split_dataset(dataset):
+def split_dataset(dataset: Dataset):
     '''By Maria.
     A function that allows for the splitting of an ImageDataset object into subsets that can be used for training, validation, testing. Splitting is done according
     to preset proportions (8:1:1). Splitting is done using randomization, but with a fixed seed for reproducible results (elements of the dataset are shuffled,

### 2023-01-03 paper/paper.tex
Update paper.tex

@@
 
-\title{AICS project working title}
+\title{\textit{A young boy with a tree made of trees}: \\ Domain Adaptation of an LSTM-based Image Caption Generation Model}
 
@@
 \begin{abstract}
-This document is a supplement to the general instructions for *ACL authors. It contains instructions for using the \LaTeX{} style files for ACL conferences. 
-The document itself conforms to its own specifications, and is therefore an example of what your manuscript should look like.
-These instructions should be used both for papers submitted for review and for final versions of accepted papers.
+Here goes the abstract for the paper.
 \end{abstract}
@@
 
-These instructions are for authors submitting papers to *ACL conferences using \LaTeX.
-They are not self-contained.
-All authors must follow the general instructions for *ACL proceedings,\footnote{\url{http://acl-org.github.io/ACLPUB/formatting.html}} and this document contains additional instructions for the \LaTeX{} style files.
+Here we will have the introduction: some general talk, a little summary of the background reading (do we need more reading?), our questions that we want answered.
 
-The templates include the \LaTeX{} source of this document (\texttt{acl.tex}),
-the \LaTeX{} style file used to format it (\texttt{acl.sty}),
-an ACL bibliography style (\texttt{acl\_natbib.bst}),
-an example bibliography (\texttt{custom.bib}),
-and the bibliography for the ACL Anthology (\texttt{anthology.bib}).
+\section{Materials and Methods}
 
-\section{Engines}
+Here we need to describe the databases that we used and the code that we used, and what we added on our own. Then we need to describe how we went about evaluation.
 
-To produce a PDF file, pdf\LaTeX{} is strongly recommended (over original \LaTeX{} plus dvips+ps2pdf or dvipdf).
-Xe\LaTeX{} also produces PDF files, and is especially suitable for text in non-Latin scripts.
+\section{Results}
 
-\section{Preamble}
+Here we should summarize our results, in terms of:
+\begin{enumerate}
+    \item where did the performance exceed the best prior performance (we got BEST checkpoints?)
+    \item which models actually generated reasonable captions? where did they start falling apart?
+    \item subjectively, which captions were best? include the questionnaire results here or subsequently
+\end{enumerate}
+Include figures and tables as much as we can/as much as is reasonable.
 
-The first line of the file must be
-\begin{quote}
-\begin{verbatim}
-\documentclass[11pt]{article}
-\end{verbatim}
-\end{quote}
+\section{Discussion}
 
-To load the style file in the review version:
-\begin{quote}
-\begin{verbatim}
-\usepackage[review]{acl}
-\end{verbatim}
-\end{quote}
-For the final version, omit the \verb|review| option:
-\begin{quote}
-\begin{verbatim}
-\usepackage{acl}
-\end{verbatim}
-\end{quote}
+Here we compare our results to our initial expectations and decide how they answered questions, we also contrast them with prior work (if there is any).
+Highlight why what we did was relevant to the field.
 
-To use Times Roman, put the following in the preamble:
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\end{verbatim}
-\end{quote}
-(Alternatives like txfonts or newtx are also acceptable.)
+\section{Conclusions and further work}
 
-Please see the \LaTeX{} source of this document for comments on other packages that may be useful.
-
-Set the title and author using \verb|\title| and \verb|\author|.
-Within the author list, format multiple authors using \verb|\and| and \verb|\And| and \verb|\AND|; please see the \LaTeX{} source for examples.
-
-By default, the box containing the title and author names is set to the minimum of 5 cm. If you need more space, include the following in the preamble:
-\begin{quote}
-\begin{verbatim}
-\setlength\titlebox{<dim>}
-\end{verbatim}
-\end{quote}
-where \verb|<dim>| is replaced with a length. Do not set this length smaller than 5 cm.
-
-\section{Document Body}
-
-\subsection{Footnotes}
-
-Footnotes are inserted with the \verb|\footnote| command.\footnote{This is a footnote.}
-
-\subsection{Tables and figures}
-
-See Table~\ref{tab:accents} for an example of a table and its caption.
-\textbf{Do not override the default caption sizes.}
+Here we should summarize what we have done and suggest what more could be done on this topic.
 
@@
 \hline
-\textbf{Command} & \textbf{Output}\\
+\textbf{This is} & \textbf{a table}\\
 \hline
-\verb|{\"a}| & {\"a} \\
-\verb|{\^e}| & {\^e} \\
-\verb|{\`i}| & {\`i} \\ 
-\verb|{\.I}| & {\.I} \\ 
-\verb|{\o}| & {\o} \\
-\verb|{\'u}| & {\'u}  \\ 
-\verb|{\aa}| & {\aa}  \\\hline
+a table & with things \\
+and even & more things  \\\hline
 \end{tabular}
-\begin{tabular}{lc}
-\hline
-\textbf{Command} & \textbf{Output}\\
-\hline
-\verb|{\c c}| & {\c c} \\ 
-\verb|{\u g}| & {\u g} \\ 
-\verb|{\l}| & {\l} \\ 
-\verb|{\~n}| & {\~n} \\ 
-\verb|{\H o}| & {\H o} \\ 
-\verb|{\v r}| & {\v r} \\ 
-\verb|{\ss}| & {\ss} \\
-\hline
-\end{tabular}
-\caption{Example commands for accented characters, to be used in, \emph{e.g.}, Bib\TeX{} entries.}
+\caption{Example table.}
 \label{tab:accents}
 \end{table}
-
-\subsection{Hyperlinks}
-
-Users of older versions of \LaTeX{} may encounter the following error during compilation: 
-\begin{quote}
-\tt\verb|\pdfendlink| ended up in different nesting level than \verb|\pdfstartlink|.
-\end{quote}
-This happens when pdf\LaTeX{} is used and a citation splits across a page boundary. The best way to fix this is to upgrade \LaTeX{} to 2018-12-01 or later.
-
-\subsection{Citations}
-
-\begin{table*}
-\centering
-\begin{tabular}{lll}
-\hline
-\textbf{Output} & \textbf{natbib command} & \textbf{Old ACL-style command}\\
-\hline
-\citep{Gusfield:97} & \verb|\citep| & \verb|\cite| \\
-\citealp{Gusfield:97} & \verb|\citealp| & no equivalent \\
-\citet{Gusfield:97} & \verb|\citet| & \verb|\newcite| \\
-\citeyearpar{Gusfield:97} & \verb|\citeyearpar| & \verb|\shortcite| \\
-\hline
-\end{tabular}
-\caption{\label{citation-guide}
-Citation commands supported by the style file.
-The style is based on the natbib package and supports all natbib citation commands.
-It also supports commands defined in previous ACL style files for compatibility.
-}
-\end{table*}
-
-Table~\ref{citation-guide} shows the syntax supported by the style files.
-We encourage you to use the natbib styles.
-You can use the command \verb|\citet| (cite in text) to get ``author (year)'' citations, like this citation to a paper by \citet{Gusfield:97}.
-You can use the command \verb|\citep| (cite in parentheses) to get ``(author, year)'' citations \citep{Gusfield:97}.
-You can use the command \verb|\citealp| (alternative cite without parentheses) to get ``author, year'' citations, which is useful for using citations within parentheses (e.g. \citealp{Gusfield:97}).
-
-\subsection{References}
-
-\nocite{Ando2005,andrew2007scalable,rasooli-tetrault-2015}
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly follow the American Psychological Association format.
-If your own bib file is named \texttt{custom.bib}, then placing the following before any appendices in your \LaTeX{} file will generate the references section for you:
-\begin{quote}
-\begin{verbatim}
-\bibliography{custom}
-\end{verbatim}
-\end{quote}
-
-You can obtain the complete ACL Anthology as a Bib\TeX{} file from \url{https://aclweb.org/anthology/anthology.bib.gz}.
-To include both the Anthology and your own .bib file, use the following instead of the above.
-\begin{quote}
-\begin{verbatim}
-\bibliography{anthology,custom}
-\end{verbatim}
-\end{quote}
-
-Please see Section~\ref{sec:bibtex} for information on preparing Bib\TeX{} files.
-
-\subsection{Appendices}
-
-Use \verb|\appendix| before any appendix section to switch the section numbering over to letters. See Appendix~\ref{sec:appendix} for an example.
-
-\section{Bib\TeX{} Files}
-\label{sec:bibtex}
-
-Unicode cannot be used in Bib\TeX{} entries, and some ways of typing special characters can disrupt Bib\TeX's alphabetization.
-The recommended way of typing special characters is shown in Table~\ref{tab:accents}.
-
-Please ensure that Bib\TeX{} records contain DOIs or URLs when possible, and for all the ACL materials that you reference.
-Use the \verb|doi| field for DOIs and the \verb|url| field for URLs.
-If a Bib\TeX{} entry has a URL or DOI field, the paper title in the references section will appear as a hyperlink to the paper, using the hyperref \LaTeX{} package.
-
-\section*{Acknowledgements}
-
-This document has been adapted
-by Steven Bethard, Ryan Cotterell and Rui Yan
-from the instructions for earlier ACL and NAACL proceedings, including those for 
-ACL 2019 by Douwe Kiela and Ivan Vuli\'{c},
-NAACL 2019 by Stephanie Lukin and Alla Roskovskaya, 
-ACL 2018 by Shay Cohen, Kevin Gimpel, and Wei Lu, 
-NAACL 2018 by Margaret Mitchell and Stephanie Lukin,
-Bib\TeX{} suggestions for (NA)ACL 2017/2018 from Jason Eisner,
-ACL 2017 by Dan Gildea and Min-Yen Kan, 
-NAACL 2017 by Margaret Mitchell, 
-ACL 2012 by Maggie Li and Michael White, 
-ACL 2010 by Jing-Shin Chang and Philipp Koehn, 
-ACL 2008 by Johanna D. Moore, Simone Teufel, James Allan, and Sadaoki Furui, 
-ACL 2005 by Hwee Tou Ng and Kemal Oflazer, 
-ACL 2002 by Eugene Charniak and Dekang Lin, 
-and earlier ACL and EACL formats written by several people, including
-John Chen, Henry S. Thompson and Donald Walker.
-Additional elements were taken from the formatting instructions of the \emph{International Joint Conference on Artificial Intelligence} and the \emph{Conference on Computer Vision and Pattern Recognition}.
 
@@
 
-This is an appendix.
+Here we should link our repository and also make sure that it is well-organized and has proper READMEs.
 

### 2023-01-03 paper/paper.tex
Abstract, part of the bibliography and intro

@@
 \begin{abstract}
-Here goes the abstract for the paper.
+
+Within this project we aim to explore how finetuning a CNN and LSTM-based image captioning model on a new dataset influences its performance on both the old and new data, along with what role some of the parameters, such as finetuning data size or the prevalence of out-of-vocabulary tokens in captions play in the process.
+We discover that differences in vocabulary play a major role and that providing too much fine-tuning data can cause the whole system to fail.
+We conclude that domain adaptation can be difficult to carry out, but perhaps worth attempting in certain cases.\footnote{MS}
+
 \end{abstract}
@@
 
-Here we will have the introduction: some general talk, a little summary of the background reading (do we need more reading?), our questions that we want answered.
+\textit{Here we will have the introduction: some general talk, a little summary of the background reading (do we need more reading?), our questions that we want answered.}
+
+Automated image caption generation is an intersectional topic, combining advances in the fields of computer vision and natural language processing.
+It is not only an interesting and challenging task, but also one with useful real-life implementations, such as assisting people with vision impairments in a variety of ways.
+It is therefore quite relevant to explore a variety of issues that can be encountered in the field, in order to improve the performance of potential models and increase their relevance in practical applications.\footnote{MS}
+
+One potential issue with image captioning systems is that of domain adaptation, meaning adapting a trained system to somewhat different data.
+While it is not uncommon to encounter e.g.
+pre-trained image processing components in image caption generation models, it would appear that the general consensus is to train a new captioning model for every new dataset.
+This, however, may lead to an unnecessary waste of resources, especially if large amounts of data are used for training; another potential issue is there not always being enough data to train a model from scratch.
+An alternative in that case would be domain adaptation.\footnote{MS}
+
+Some more lightweight, but well-performing image captioning architectures include ones with a CNN-based image processing component and LSTM-based language generation component; their performance can be further increased by incorporating some form of attention, which allows for a more precise identification of what is relevant to the caption generation at what point.
+The particular model used in this project is based on \cite{xu2015attend}, with the implementation by \cite{ilinykh}.
+In the model, the encoder is a convolutional network, from which representations of elements of the picture - feature vectors - are obtained.
+The decoder is an LSTM network with attention, determining what parts of the image are the most relevant for generating a given element of the caption (ADD WHICH KIND OF ATTENTION?).\footnote{MS}
+
+Another way of employing attention in a similar model is presented in \cite{lu16}, where a special "visual sentinel" helps determine when to attend to the data from the image and when to the information from the language model, as not all the words are equally represented in an image - for instance, articles or other determiners, particles, etc., stem more from the requirements of a given grammatical structure or content word rather than the scene that is being described.
+However, since this project did not delve into the contribution of each of these elements in caption generation, it was found to be superfluous in our particular case, even though it can lead to improved performance.\footnote{MS}
 
@@
 Here we need to describe the databases that we used and the code that we used, and what we added on our own. Then we need to describe how we went about evaluation.
+
+\subsection{The image captioning model}
+\subsection{Flickr8k}
+\subsection{imageCLEF}
+\subsection{Human judgement questionnaire}
 

### 2023-01-03 paper/paper.tex
Updating the bibliography, finishing the introduction

@@
 
+The topic of domain adaptation in caption generation itself does not appear to be widely studied.
+However, \cite{hesselSW15} touche upon it in their paper on the importance of the language model versus the image processing component, where they prove that a good language model can produce decent captions even with poor image input.
+However, what the authors actually discuss is only fine-tuning the image recognition element of the model.
+While that is naturally helpful, especially since models of that kind require massive amounts of training data, it does not fully address the issue of fine-tuning a whole image caption generation model, and, in fact, utilizing pre-trained CNN components has become rather common.\footnote{MS}
+
+While all of the aforementioned papers seem to approach this issue as "generation from visual input," \cite{bernardi16} describe other approaches to the issue of image caption generation, showing that this is simply one of the possible routes.
+Some shortcomings of this approach that they name include the inability of the system to generate captions for visual input where the system cannot recognize any salient elements.
+They also tackle the issue of evaluation, listing the variety of measures that can be used to that end when it comes to caption generation (as, naturally, simpler ones like accuracy would not be a good match); they admit that human judgement or measures that mimic human judgement are key when it comes to evaluating automatically generated captions.\footnote{MS}
+
+Given how little information we could fine on the fine-tuning of CNN+LSTM image captioning models, within this project we would like to explore the extent to which a trained model of such architecture can be adapted to another domain and how some of the hyperparameters of that process can influence the outcome.
+We expect this process to be effective to an extent, but it may also break the model at some point.
+We hope for the fine-tuned model to perform better on the images from the new domain than the non-fine-tuned one.\footnote{MS}
+
+In Section 2 we describe the materials and methods used to address the aforementioned questions, such as the architecture of the model, the datasets, and the evaluation methods.
+In Section 3 we present the results of our inquiry.
+In Section 4 we discuss how they answer our questions and how they relate to previous work.
+Finally, in Section 5 we offer conclusions and suggestions for future projects on this topic.\footnote{MS}
+
 \section{Materials and Methods}

### 2023-01-04 paper/paper.tex
Splitting the writing

@@
 
-\subsection{The image captioning model}
-\subsection{Flickr8k}
-\subsection{imageCLEF}
-\subsection{Human judgement questionnaire}
+\subsection{The image captioning model \textbf{DK}}
+\subsection{Flickr8k \textbf{MS or DK}}
+\subsection{imageCLEF \textbf{MS or DK}}
+\subsection{Human judgement questionnaire \textbf{MS}}
 
-\section{Results}
+\section{Results \textbf{MS and DK?}}
 
@@
 \begin{enumerate}
-    \item where did the performance exceed the best prior performance (we got BEST checkpoints?)
-    \item which models actually generated reasonable captions? where did they start falling apart?
-    \item subjectively, which captions were best? include the questionnaire results here or subsequently
+    \item where did the performance exceed the best prior performance (we got BEST checkpoints?) \textbf{DK}
+    \item which models actually generated reasonable captions? where did they start falling apart? \textbf{MS or DK}
+    \item subjectively, which captions were best? include the questionnaire results here or subsequently \textbf{MS}
 \end{enumerate}
@@
 
-\section{Discussion}
+\section{Discussion \textbf{MS and/or DK}}
 
@@
 
-\section{Conclusions and further work}
+\section{Conclusions and further work \textbf{MS and/or DK}}
 

### 2023-01-05 paper/paper.tex
Methods: questionnaire

@@
 We expect this process to be effective to an extent, but it may also break the model at some point.
-We hope for the fine-tuned model to perform better on the images from the new domain than the non-fine-tuned one.\footnote{MS}
+We hope for the fine-tuned model to perform better on the images from the new domain than the non-fine-tuned one, which we aim to evaluate using human judgement.
+Throughout the project, Dominik focused on the influence of hyperparameters, while Maria on the evaluation, with both authors inventing ways to adapt the model to their needs and identify and fix potential issues.\footnote{MS}
 
@@
 
-Here we need to describe the databases that we used and the code that we used, and what we added on our own. Then we need to describe how we went about evaluation.
+\textit{Here we need to describe the databases that we used and the code that we used, and what we added on our own. Then we need to describe how we went about evaluation.}
+
+Within this section we will describe the resources that we used in the project.
+We will specify the characteristics of the image captioning model, the two datasets that were used to train or fine-tune it, and the questionnaire we utilized to elicit human judgements of the best captions.\footnote{MS}
 
 \subsection{The image captioning model \textbf{DK}}
+\subsection{The code \textbf{MS and DK}}
 \subsection{Flickr8k \textbf{MS or DK}}
@@
 
+Having produced results (captions) using the model and the data described in the previous sections, we decided to test which of the plausible-sounding captions were the best match for the images.
+The only models that were somewhat consistent producing coherent captions were the non-fine-tuned (original) model and the "100" training data size models (with 4 variations of UNK filter: 0.1, 0.15, 0.2, no filter).
+In order to collect judgements, 20 images from 3 test sets each were collected, together with the captions generated by the aforementioned models.
+The test sets were: unfiltered imageCLEF, filtered (max 10\% UNK tokens) imageCLEF, and Flickr8k.
+The questionnaire was constructed and hosted using Google Forms.
+
+The questionnaire itself consisted of five sections.
+In the first one, a modicum of personal information, such as age, gender, education, native language was collected.
+The following three sections were constructed the same way, and each of them corresponded to a different test set.
+In each of those sections, 20 images, along with the corresponding captions were presented.
+The order of the captions for each image was randomized to avoid bias.
+The participants were instructed to select one caption for each image which they thought was the best fit or the best description of the image.
+At the end of every section the participants had an opportunity to voice their general thoughts on the captions.
+Finally, in the last section the participants had the chance to include their final thoughts and opinions before submitting their answers.
+
+The form was subsequently distributed online, and the responses were later processed and a summary will be presented in the Results section.
+
 \section{Results \textbf{MS and DK?}}
 
-Here we should summarize our results, in terms of:
+\textit{Here we should summarize our results, in terms of:
 \begin{enumerate}
@@
 \end{enumerate}
-Include figures and tables as much as we can/as much as is reasonable.
+Include figures and tables as much as we can/as much as is reasonable.}
+
+\subsection{Hyperparameters \textbf{DK}}
+
+\subsection{Caption Evaluation \textbf{MS}}
 

### 2023-01-05 paper/paper.tex
paper resources

@@
 
-Given how little information we could fine on the fine-tuning of CNN+LSTM image captioning models, within this project we would like to explore the extent to which a trained model of such architecture can be adapted to another domain and how some of the hyperparameters of that process can influence the outcome.
+Given how little information we could find on the fine-tuning of CNN+LSTM image captioning models, within this project we would like to explore the extent to which a trained model of such architecture can be adapted to another domain and how some of the hyperparameters of that process can influence the outcome.
 We expect this process to be effective to an extent, but it may also break the model at some point.
@@
 
-Here we compare our results to our initial expectations and decide how they answered questions, we also contrast them with prior work (if there is any).
-Highlight why what we did was relevant to the field.
+\textit{Here we compare our results to our initial expectations and decide how they answered questions, we also contrast them with prior work (if there is any).
+Highlight why what we did was relevant to the field.}
 
@@
 
-Here we should summarize what we have done and suggest what more could be done on this topic.
+\textit{Here we should summarize what we have done and suggest what more could be done on this topic.}
 

### 2023-01-05 paper/paper.tex
imageCLEF, some results

--- 2023-01-05 paper/paper.tex
@@
 \usepackage{microtype}
+\usepackage{float}
 
@@
 \subsection{The code \textbf{MS and DK}}
-\subsection{Flickr8k \textbf{MS or DK}}
-\subsection{imageCLEF \textbf{MS or DK}}
+
+\textit{Write about the division into files, a bit what's what, what we added and what was there.
+I did the evaluation part.
+I doubt we need to describe it in much detail, can always say that the readme has instructions and all the elements are documented, but I'd still have a bit of a summary of what we did here.}
+
+The testing and evaluation part was conducted in a variety of Jupyter Notebook files.
+While the initial plan was to contain all the information in one file; unfortunately, due to GitHub limitations, that Notebook would have been too large to upload due to the number of generated ans printed image and caption pairs.
+Therefore, all training data sized but 2k and 5k were tested in separate notebooks using the same methods, adapted from \cite{ilinykh}, stored in a separate file, and imported into the notebook.
+The captions were generated using beam search and then presented together with the mappings of the attention on the image.
+\footnote{MS}
+
+Other than relatively minor fixes that were required to make the code run, we decided to manually exclude the UNK token from the generation process by enforcing the lowest possible probability on it in the beam search.
+While this meant that we could not evaluate to what extent the models favored generating that token, it did allow us to obtain captions that sounded more natural, which was paramount for the subsequent human evaluation.\footnote{MS}
+
+\subsection{Flickr8k \textbf{DK}}
+\subsection{imageCLEF \textbf{MS}}
+
+The second dataset used in the project was the IAPR TC-12 Benchmark, called imageCLEF in this paper.
+This dataset consists of 20,000 images from a variety of real life situations.
+Each image is paired with a description consisting of a number of consecutive utterances specifying what the image is depicting \cite{imageclef}.
+The assumption made in this project is that the first fragment of the description contains the most salient elements.
+Unlike Flickr8k, imageCLEF contains many more images that do not include people or animals.
+Additionally, the way the descriptions are formulated is also different from Flickr8k.
+
 \subsection{Human judgement questionnaire \textbf{MS}}
@@
 Finally, in the last section the participants had the chance to include their final thoughts and opinions before submitting their answers.
+What is important to mention is that sometimes the model could not generate a caption for a given image, or the captions were the same across multiple models.
+Luckily, there were always at least two different captions.
+Nevertheless, this does make the interpretation of the results somewhat more difficult.
 
@@
 
+As described previously in the Materials and Methods section, the only models that consistently produced captions for almost all test images were the original and the "100" models.
+The "200" models still generated most of the captions, but many of them started being rather nonsensical or non-grammatical.
+This process continued in the "500" and "1000" models with fewer and fewer captions being generated successfully, and those generated ones being absolutely nonsensical.
+Finally in the "2k" and "5k" models most captions could not be generated.\footnote{MS} \textit{This part can be moved somewhere else if it fits better for your hyperparameters section.}
+
+The questionnaire was filled out by 6 people.
+While we would like a larger sample size, we believe that we can still draw some preliminary conclusions from their answers, and that it is sufficient taking the scope of the project into account.
+It is also likely that the length of the questionnaire could have put off other potential respondents.
+The people who filled the questionnaire out ranged between 25 and 31 years of age.
+Three of them identified as female, two as male, one selected "other" as a response in that option.
+Half of the respondents had an MA/MSc degree or equivalent, two a BA/BSc or equivalent, and one selected secondary school education as the highest completed level of education.
+None of the people who filled the questionnaire out were native speakers of English.
+More detailed data about their native language was not collected.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{1cm}|p{4.5cm}|}
+\hline \bf Image & \bf Best model (\% responses) \\ \hline
+15551 & 100\_0.1 (50\%)  \\
+20144 & original (50\%)  \\
+20343 & 100\_0.1 (33.3\%) or 100\_1.0 (33.3\%)  \\
+12761 & original (33.3\%) or 100\_1.0 (33.3\%)  \\
+22381 & 100\_0.1 (83.3\%) \\
+10821 & 100\_0.1 (100\%) \\
+18448 & original (100\%) \\
+17173 & original (66.7\%) \\
+13152 & original (66.7\%)  \\
+20272 & original (100\%)  \\
+11310 & 100\_0.1 (66.7\%)  \\
+14183 & 100\_1.0 (50\%)  \\
+10622 & original (50\%)  \\
+13123 & original (100\%)   \\
+19181 & 100\_1.0 (83.3\%) \\
+16920 & 100\_0.1 (100\%) \\
+11308 & original (50\%) \\
+14016 & original (83.3\%) \\
+21029 & original (100\%) \\
+18355 & original (33.3\%) or 100\_0.15 (33.3\%) or 100\_0.2 (33.3\%) \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{unfiltered-table} Best captions per image in the unfiltered imageCLEF test set. }
+\end{table}
+
 \section{Discussion \textbf{MS and/or DK}}
 
-Here we compare our results to our initial expectations and decide how they answered questions, we also contrast them with prior work (if there is any).
-Highlight why what we did was relevant to the field.
+\textit{Here we compare our results to our initial expectations and decide how they answered questions, we also contrast them with prior work (if there is any).
+Highlight why what we did was relevant to the field.}
 
@@
 
-Here we should link our repository and also make sure that it is well-organized and has proper READMEs.
+\textit{Here we should link our repository and also make sure that it is well-organized and has proper READMEs.}
 

### 2023-01-05 paper/paper.tex
Adding tables

--- 2023-01-05 paper/paper.tex
@@
 
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{1cm}|p{4.5cm}|}
+\hline \bf Image & \bf Best model (\% responses) \\ \hline
+40416 & original (83.3\%)  \\
+39158 & original (50\%)  \\
+25053 & original (83.3\%)  \\
+30620 & 100\_1.0 (66.7\%)  \\
+32397 & 100\_0.2 (66.7\%) \\
+38937 & original (50\%) \\
+39005 & 100\_0.1 (33.3\%) or 100\_0.15 (33.3\%) or 100\_0.2 (33.3\%) \\
+40120 & original (66.7\%) \\
+23588 & 100\_1.0 (83.3\%)  \\
+39472 & 100\_0.1 (66.7\%)  \\
+30138 & original (50\%)  \\
+40202 & 100\_0.1 (83.3\%)  \\
+35895 & original (66.7\%)  \\
+32663 & 100\_0.2 (83.3\%)   \\
+30705 & original (83.3\%) \\
+38081 & 100\_0.15 (66.7\%) \\
+31571 & original (100\%) \\
+35858 & original (66.7\%) \\
+37836 & 100\_0.1 (33.3\%) or 100\_0.15 (33.3\%) \\
+39239 & 100\_0.15, 100\_0.2, and 100\_1.0 (50\%)\footnote{These models generated identical captions.} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{filtered-table} Best captions per image in the filtered imageCLEF test set. }
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{2cm}|p{3.5cm}|}
+\hline \bf Image & \bf Best model (\% responses) \\ \hline
+2116444946\_ 1f5d1fe5d1 & original (100\%)  \\
+2316097768\_ ef662f444b & original (100\%)  \\
+2439384468\_ 58934deab6 & original (66.7\%)  \\
+2112921744\_ 92bf706805 & 100\_0.2 (50\%)  \\
+2392460773\_ 2aa01eb340 & original (100\%) \\
+2434006663\_ 207a284cec & original (100\%) \\
+2308256827\_ 3c0a7d514d & original (100\%) \\
+2111360187\_ d2505437b7 & original (66.7\%) \\
+2271671533\_ 7538ccd556 & original (66.7\%)  \\
+2328616978\_ fb21be2b87 & original (100\%)  \\
+2456907314\_ 49bc4591c4 & original (66.7\%)  \\
+2229179070\_ dc8ea8582e & 100\_0.1 (100\%)  \\
+2279980395\_ 989d48ae72 & 100\_0.1 (88.3\%)  \\
+2393971707\_ bce01ae754 & original (83.3\%)   \\
+211277478\_ 7d43aaee09 & 100\_0.1 (50\%) \\
+2337919839\_ df83827fa0 & original (83.3\%) \\
+2447035752\_ 415f4bb152 & original (50\%) \\
+23445819\_ 3a458716c1 & 100\_0.2 (66.7\%) \\
+2448210587\_ 9fe7ea5f42 & original (33.3\%) \\
+2445654384\_ 4ee3e486e1 & 100\_0.1 (50\%) or 100\_0.2 (50\%)\\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{flickr-table} Best captions per image in the Flickr8k test set. }
+\end{table}
+
 \section{Discussion \textbf{MS and/or DK}}

### 2023-01-05 paper/paper.tex
Results (questionnaire)

--- 2023-01-05 paper/paper.tex
@@
     \item which models actually generated reasonable captions? where did they start falling apart? \textbf{MS or DK}
-    \item subjectively, which captions were best? include the questionnaire results here or subsequently \textbf{MS}
 \end{enumerate}
@@
 
+While the captions themselves can be seen in the Jupyter Notebook files, the aim of the questionnaire was to reveal which model or models performed the best on which test set.
+Therefore, the results presented here are divided between the sets and include only the information on which model's caption was deemed the best, as well as how many people voted for it.
+Table 1 contains the information for the unfiltered imageCLEF set, Table 2 - for the filtered one, and Table 3 for Flickr8k.
+Image names are the file names without the .jpg extension.
+In some cases more than one model's caption received the same amount of votes, or the winning caption was generated by more than one model.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{1cm}|p{4.5cm}|}
+\hline \bf Image & \bf Best model (\% responses) \\ \hline
+40416 & original (83.3\%)  \\
+39158 & original (50\%)  \\
+25053 & original (83.3\%)  \\
+30620 & 100\_1.0 (66.7\%)  \\
+32397 & 100\_0.2 (66.7\%) \\
+38937 & original (50\%) \\
+39005 & 100\_0.1 (33.3\%) or 100\_0.15 (33.3\%) or 100\_0.2 (33.3\%) \\
+40120 & original (66.7\%) \\
+23588 & 100\_1.0 (83.3\%)  \\
+39472 & 100\_0.1 (66.7\%)  \\
+30138 & original (50\%)  \\
+40202 & 100\_0.1 (83.3\%)  \\
+35895 & original (66.7\%)  \\
+32663 & 100\_0.2 (83.3\%)   \\
+30705 & original (83.3\%) \\
+38081 & 100\_0.15 (66.7\%) \\
+31571 & original (100\%) \\
+35858 & original (66.7\%) \\
+37836 & 100\_0.1 (33.3\%) or 100\_0.15 (33.3\%) \\
+39239 & 100\_0.15, 100\_0.2, and 100\_1.0 (50\%)\footnote{These models generated identical captions.} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{filtered-table} Best captions per image in the filtered imageCLEF test set. }
+\end{table}
+
+As can be noticed in Table 1, the majority of the "good" captions for the unfiltered imageCLEF set were generated by the original, non-fine-tuned model, or by the one with a 10\% filter.
+Some captions from the model fine-tuned on nonfiltered data turned out to have been rated high as well.
+Finally, other models' captions were rated as good when they overlapped with the original model's caption.
+Overall, while the judgements are split rather evenly, the original model seems to generate better captions for this test set.
+It is worth pointing out that relatively few captions were unanimously voted best.
+
+When it comes to the filtered imageCLEF test set (Table 2), where a maximum of 10\% tokens in the original caption could be unknown tokens, the spectrum of what models' captions were considered good is even wider.
+There was only one case where one caption was voted best by all the participants.
+Half of the best captions were generated by the original model, and half by various fine-tuned ones, with none of them being a clear front-runner in this case.
+
+Finally, when it comes to the original Flickr8k-based test set, as seen in Table 3, the original model without fine-tuning performed best, although there were instances where the 10\% or 20\% fine-tuned models generated better captions.
+Many of the original model's captions were also chosen unanimously.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{2.5cm}|p{3cm}|}
+\hline \bf Image & \bf Best model (\% responses) \\ \hline
+2116444946[...] & original (100\%)  \\
+2316097768[...] & original (100\%)  \\
+2439384468[...] & original (66.7\%)  \\
+2112921744[...] & 100\_0.2 (50\%)  \\
+2392460773[...] & original (100\%) \\
+2434006663[...] & original (100\%) \\
+2308256827[...] & original (100\%) \\
+2111360187[...] & original (66.7\%) \\
+2271671533[...] & original (66.7\%)  \\
+2328616978[...] & original (100\%)  \\
+2456907314[...] & original (66.7\%)  \\
+2229179070[...] & 100\_0.1 (100\%)  \\
+2279980395[...] & 100\_0.1 (88.3\%)  \\
+2393971707[...] & original (83.3\%)   \\
+211277478[...] & 100\_0.1 (50\%) \\
+2337919839[...] & original (83.3\%) \\
+2447035752[...] & original (50\%) \\
+23445819[...] & 100\_0.2 (66.7\%) \\
+2448210587[...] & original (33.3\%) \\
+2445654384[...] & 100\_0.1 (50\%) or 100\_0.2 (50\%)\\
+\hline
+\end{tabular}
+\end{center}
+\caption{\label{flickr-table} Best captions per image in the Flickr8k test set. }
+\end{table}
+
+When it comes to the comments left by the respondents after each section, they were not very positive.
+The captions in the first content section were judged as "unfitting," "strange," "ungrammatical," "inaccurate," "nonsensical," or "confusing," with some of the respondents admitting that the mismatch between the image and captions made them anxious or made their "brain hurt." For the second content section, some people considered the captions better, and some worse than previously, and the feeling of anxiety persisted here too.
+In addition, half of the respondents pointed out that the captions tend to include people in the descriptions even when there are no people in the images.
+It seems that the impressions of the captions in the last content section were better: while people said that they were still "strange," "weird," or "confusing," they also admitted that the verbs in the sentences tended to be accurate descriptions of the events in the images, and that they were overall "somewhat accurate." The person who expressed their feelings of anxiety previously now said that they "started to get used to them," which could also be a consequence of there being a smaller mis-match between the images and the captions.
+
+In the final remarks many of the respondents expressed their dissatisfaction with the overall quality of the captions, leaving comments such as "r u b b i s h" or "my brain melted." Two people mentioned again how the captions would include people despite there not being any in the images.
+One person said that the simpler captions or images were clear, while the more complex ones were problematic or mismatched.
+Another person said that these "didn't sound like human-written captions," and yet another that they oftentimes "missed the essence of the picure."
+
+Overall, the original model performed best on the original test set.
+The most captions generated by fine-tuned models were selected as best for the filtered imageCLEF test set.
+The unfiltered test set was also fairly balanced between the fine-tuned and original models.
+While for the unfiltered data the unfiltered and 10\% fine-tuned models can be identified as the best out of the fine-tuned ones, the same is impossible to determine for the filtered test set.
+From these results it can be concluded that while fine-tuning does improve the quality of the captions somewhat (both for imageCLEF and Flickr8k), it does not have that effect for every image.
+Judging by the open responses from the participants, the captions generated for Flickr8k were the least wrong, but their quality was still low, and the ones for imageCLEF were even worse.
+
+Clearly, while some degree of improvement can be obtained by this kind of fine-tuning, it is far from perfect and the captions rarely match the images.
+This is likely due to the disparity between the kinds of images in the two sets.
+Another clue to that being a major issue is the models' tendency to - as pointed out by the respondents - insert people into captions when there were none in the original image.
+A large portion of the captions in the Flickr8k dataset are constructed that way, and it must be one of the things that the model learned to try to identify in the image at all cost and feature in the caption, and that the fine-tuning did not help mitigate.
+
+Finally, it is worth keeping in mind that some of the apparent weirdness of the captions may stem from the relatively small vocabulary that the models had, which was likely lacking the proper words to describe certain elements in the images, even when the ones whose captions contained many new words were filtered out.
+
 \section{Discussion \textbf{MS and/or DK}}

### 2023-01-05 paper/paper.tex
Conclusions

@@
 
-\section{Conclusions and further work \textbf{MS and/or DK}}
+\section{Conclusions and further work \textbf{MS}}
 
 \textit{Here we should summarize what we have done and suggest what more could be done on this topic.}
+
+Throughout this project we have explored and tested how a CNN and LSTM-based image caption generation model could be fine-tuned, or adapted to a new domain, as represented by a somewhat different image-caption dataset.
+We have identified points that can be especially problematic, such as differences in the vocabulary and the subsequent generation of UNK tokens in the captions.
+We have determined that if the domains are too different and too much fine-tuning data is fed to the model, the generation will start to completely fail or focus only on a very narrow range of words, producing very ungrammatical captions.
+We have also collected human judgements and determined that while fine-tuning does slighly increase the quality of the captions on the new dataset, all of them, regardless of the model, were still considered weird, mismatched, or unnatural, and that the structure of the sentences in the original dataset may have a large influence on how they are generated even after fine-tuning (e.g.
+trying to feature people in every caption).
+
+While we have addressed the issues that we raised in the introduction, we were not able to explore all the possible variables in domain adaptation of models of this kind, as it was beyond the scope of this project.
+However, we are certain that it would be very interesting to see how other hyperparameters than just the training data size can influence the outcome, and what the interplay between them could be.
+It would also be fascinating to see similar research pertaining to other model architectures, especially trying to identify the optimal hyperparameters for fine-tuning transformer-based image caption generation models.
+Finally, it would be interesting to explore other potential ways of dealing with the vocabulary mismatch between different datasets.
 

### 2023-01-06 paper/paper.tex
Paper updates

@@
 We discover that differences in vocabulary play a major role and that providing too much fine-tuning data can cause the whole system to fail.
-We conclude that domain adaptation can be difficult to carry out, but perhaps worth attempting in certain cases.\footnote{MS}
+We conclude that domain adaptation can be difficult to carry out, but perhaps worth attempting in certain cases.
 
@@
 
-\section{Introduction}
-
-\textit{Here we will have the introduction: some general talk, a little summary of the background reading (do we need more reading?), our questions that we want answered.}
+\section{Introduction (MS)}
 
@@
 It is not only an interesting and challenging task, but also one with useful real-life implementations, such as assisting people with vision impairments in a variety of ways.
-It is therefore quite relevant to explore a variety of issues that can be encountered in the field, in order to improve the performance of potential models and increase their relevance in practical applications.\footnote{MS}
+It is therefore quite relevant to explore a variety of issues that can be encountered in the field, in order to improve the performance of potential models and increase their relevance in practical applications.
 
@@
 This, however, may lead to an unnecessary waste of resources, especially if large amounts of data are used for training; another potential issue is there not always being enough data to train a model from scratch.
-An alternative in that case would be domain adaptation.\footnote{MS}
+An alternative in that case would be domain adaptation.
 
@@
 In the model, the encoder is a convolutional network, from which representations of elements of the picture - feature vectors - are obtained.
-The decoder is an LSTM network with attention, determining what parts of the image are the most relevant for generating a given element of the caption (ADD WHICH KIND OF ATTENTION?).\footnote{MS}
+The decoder is an LSTM network with attention, determining what parts of the image are the most relevant for generating a given element of the caption.
 
 Another way of employing attention in a similar model is presented in \cite{lu16}, where a special "visual sentinel" helps determine when to attend to the data from the image and when to the information from the language model, as not all the words are equally represented in an image - for instance, articles or other determiners, particles, etc., stem more from the requirements of a given grammatical structure or content word rather than the scene that is being described.
-However, since this project did not delve into the contribution of each of these elements in caption generation, it was found to be superfluous in our particular case, even though it can lead to improved performance.\footnote{MS}
+However, since this project did not delve into the contribution of each of these elements in caption generation, it was found to be superfluous in our particular case, even though it can lead to improved performance.
 
@@
 However, what the authors actually discuss is only fine-tuning the image recognition element of the model.
-While that is naturally helpful, especially since models of that kind require massive amounts of training data, it does not fully address the issue of fine-tuning a whole image caption generation model, and, in fact, utilizing pre-trained CNN components has become rather common.\footnote{MS}
+While that is naturally helpful, especially since models of that kind require massive amounts of training data, it does not fully address the issue of fine-tuning a whole image caption generation model, and, in fact, utilizing pre-trained CNN components has become rather common.
 
@@
 Some shortcomings of this approach that they name include the inability of the system to generate captions for visual input where the system cannot recognize any salient elements.
-They also tackle the issue of evaluation, listing the variety of measures that can be used to that end when it comes to caption generation (as, naturally, simpler ones like accuracy would not be a good match); they admit that human judgement or measures that mimic human judgement are key when it comes to evaluating automatically generated captions.\footnote{MS}
+They also tackle the issue of evaluation, listing the variety of measures that can be used to that end when it comes to caption generation (as, naturally, simpler ones like accuracy would not be a good match); they admit that human judgement or measures that mimic human judgement are key when it comes to evaluating automatically generated captions.
 
@@
 We hope for the fine-tuned model to perform better on the images from the new domain than the non-fine-tuned one, which we aim to evaluate using human judgement.
-Throughout the project, Dominik focused on the influence of hyperparameters, while Maria on the evaluation, with both authors inventing ways to adapt the model to their needs and identify and fix potential issues.\footnote{MS}
+Throughout the project, Dominik focused on the influence of hyperparameters, while Maria on the evaluation, with both authors inventing ways to adapt the model to their needs and identify and fix potential issues.
 
@@
 In Section 4 we discuss how they answer our questions and how they relate to previous work.
-Finally, in Section 5 we offer conclusions and suggestions for future projects on this topic.\footnote{MS}
+Finally, in Section 5 we offer conclusions and suggestions for future projects on this topic.\footnote{(MS)}
 
-\section{Materials and Methods}
-
-\textit{Here we need to describe the databases that we used and the code that we used, and what we added on our own. Then we need to describe how we went about evaluation.}
+\section{Materials and Methods (MS and DK)}
 
 Within this section we will describe the resources that we used in the project.
-We will specify the characteristics of the image captioning model, the two datasets that were used to train or fine-tune it, and the questionnaire we utilized to elicit human judgements of the best captions.\footnote{MS}
+We will specify the characteristics of the image captioning model, the two datasets that were used to train or fine-tune it, and the questionnaire we utilized to elicit human judgements of the best captions.\footnote{(MS)}
 
-\subsection{The image captioning model \textbf{DK}}
-\subsection{The code \textbf{MS and DK}}
+\subsection{The image captioning model (DK)}
+\subsection{The code (MS and DK)}
 
@@
 The captions were generated using beam search and then presented together with the mappings of the attention on the image.
-\footnote{MS}
 
 Other than relatively minor fixes that were required to make the code run, we decided to manually exclude the UNK token from the generation process by enforcing the lowest possible probability on it in the beam search.
-While this meant that we could not evaluate to what extent the models favored generating that token, it did allow us to obtain captions that sounded more natural, which was paramount for the subsequent human evaluation.\footnote{MS}
+While this meant that we could not evaluate to what extent the models favored generating that token, it did allow us to obtain captions that sounded more natural, which was paramount for the subsequent human evaluation.\footnote{(MS)}
 
-\subsection{Flickr8k \textbf{DK}}
-\subsection{imageCLEF \textbf{MS}}
+\subsection{Flickr8k (DK)}
+\subsection{imageCLEF (MS)}
 
@@
 
-\subsection{Human judgement questionnaire \textbf{MS}}
+\subsection{Human judgement questionnaire (MS)}
 
@@
 
-\section{Results \textbf{MS and DK?}}
+\section{Results (MS and DK)}
 
@@
 
-\subsection{Hyperparameters \textbf{DK}}
+\subsection{Hyperparameters (DK)}
 
-\subsection{Caption Evaluation \textbf{MS}}
+\subsection{Caption Evaluation (MS)}
 
@@
 
-\section{Discussion \textbf{MS and/or DK}}
+\section{Discussion (DK)}
 
@@
 
-\section{Conclusions and further work \textbf{MS}}
-
-\textit{Here we should summarize what we have done and suggest what more could be done on this topic.}
+\section{Conclusions and further work (MS)}
 
@@
 
-\section{Example Appendix}
+\section{Appendix}
 \label{sec:appendix}
 
-\textit{Here we should link our repository and also make sure that it is well-organized and has proper READMEs.}
+\begin{enumerate}
+    \item The project repository with instructions: \url{https://github.com/Turtilla/aics-project}
+\end{enumerate}
+
+\section{Statement of Contribution}
+\label{sec:contribution}
+
+\textit{Here we have our statement of who did what.}
+While really detailed information on who developed what is available in the code, the code is not the only thing that we shared work on.
+Below one can find the details of who contributed what to the project.
+\begin{enumerate}
+    \item \textbf{Developing project ideas:} both Dominik and Maria.
+Selecting the initial ideas from the ones suggested in the course page mostly by Dominik, ideas about working with comicbooks and contacting researchers related to that by Maria.
+Final idea selected under the advice from Nikolai and Simon by both project authors.
+    \item \textbf{Setting up resources:} figuring out how to access the datasets used in the project by Dominik, using the existing code by Nikolai both by Maria and Dominik.
+    \item \textbf{Code:} detailed information on code contribution can be found in code documentation. 
+    \begin{enumerate}
+        \item Evaluation Jupyter Notebooks: Maria
+        \item Dataset classes: Dominik
+        \item Relation filters: Maria and code shared by Simon
+        \item Fine-tuning loop: adapted by Maria and Dominik from Nikolai's code.
+        \item Models and some preprocessing: Nikolai's code.
+        \item Caption generation and attention visualization: Nikolai's code, with changes by Maria and Dominik.
+        \item Division of the code into separate files, turning it into classes: Dominik.
+        \item Documentation: Maria with Dominik's help.
+    \end{enumerate}
+    \item \textbf{Questionnaire:} Maria.
+    \item \textbf{Evaluation:} evaluation of the outputs of the notebooks by Dominik and Maria (for the questionnaire), evaluation of questionnaire answers by Maria.
+    \item \textbf{Writing up:} both Dominik and Maria.
+Abstract by Maria (cannot put the initials there).
+While the initials signify the person who wrote most of the section, small changes by the other person are possible.
+Some sections' authorship is shared; in that case a footnote is put at the end of the part written by one person.
+    \item \textbf{Repository management:} Maria with Dominik's help.
+\end{enumerate}
 

### 2023-01-09 paper/paper.tex
typos, paper, cleanup

@@
 
-Within this project we aim to explore how finetuning a CNN and LSTM-based image captioning model on a new dataset influences its performance on both the old and new data, along with what role some of the parameters, such as finetuning data size or the prevalence of out-of-vocabulary tokens in captions play in the process.
+    Within this project we aim to explore how fine-tuning a CNN and LSTM-based image captioning model on a new dataset influences its performance on both the old and new data, along with what role some of the parameters, such as fine-tuning data size or the prevalence of out-of-vocabulary tokens in captions play in the process.
 We discover that differences in vocabulary play a major role and that providing too much fine-tuning data can cause the whole system to fail.
@@
 The topic of domain adaptation in caption generation itself does not appear to be widely studied.
-However, \cite{hesselSW15} touche upon it in their paper on the importance of the language model versus the image processing component, where they prove that a good language model can produce decent captions even with poor image input.
+However, \cite{hesselSW15} touch upon it in their paper on the importance of the language model versus the image processing component, where they prove that a good language model can produce decent captions even with poor image input.
 However, what the authors actually discuss is only fine-tuning the image recognition element of the model.
@@
 \subsection{The image captioning model (DK)}
+
+The goal of this project was to analyze how the generation of captions changes, when an already pre-trained model is fine-tuned on new data from a different domain.
+For this, we utilized code written by \cite{ilinykh} that implemented a model described by \cite{xu2015attend}.
+We didn't adapt the model at all, but only fine-tuned it with a different dataset.
+In the following paragraph, the architecture model is summarized.
+
+The model consists of an \emph{Encoder} and a \emph{Decoder}.
+The \emph{Encoder} utilizes the ResNet-101 pre-trained model to extract features of the images \cite{resnet}.
+The last two layers are removed to only include the parts important for the feature extraction.
+The \emph{Decoder} then initializes the hidden state of an LSTM with the image representation and generates the caption word by word.
+Hereby, the LSTM also makes use of an attention mechanism that weights different parts of the image higher and lower, depending on how important they are to generate a word at a certain step.
+This attention can then be visualized as a heat map on the image.
+Dropout is used to prevent overfitting to training data.
+
 \subsection{The code (MS and DK)}
@@
 
+Loading the data was done in a \emph{PyTorch Dataset} class.
+Here, it is possible to define multiple parameters for loading the data.
+First of all, only captions are loded that contain relations.
+Here, we implemented two different methods.
+One is filtering the captions based on the POS and the other is a rule-based approach taken from \cite{Ghanimifard2018}.
+Furthermore, a total number of images that should be loaded can be defined.
+This loaded samples are then divided into training (0.8), validation (0.1) and testing (0.1) sets.
+Since the imageCLEF captions consists of multiple connected captions, we included an option to either just load the first caption or concatenate all with the conjunction \emph{and}.
+Lastly, we noted that changing the dataset presents us with a lot of words that are not included in the original vocabulary.
+To deal with this, we implemented a variable filter threshold to exclude captions that contain too many unknown words (\emph{unknown filter}).
+
+The training loop was mostly taken from \cite{ilinykh}.
+The pre-trained model is hereby fine-tuned on an unseen shuffled train dataset.
+After each epoch the model is validated with a validation dataset.
+This is done using BLEU-4 score.
+Since the imageCLEF dataset only consists of one independent caption instead of five as the Flickr8k dataset, the BLEU score is calculated only with one reference caption.
+The model is trained as long as the BLEU score doesn't improve for 20 epochs; the model with the best BLEU score is saved and used for testing and evaluating.
+
 The testing and evaluation part was conducted in a variety of Jupyter Notebook files.
-While the initial plan was to contain all the information in one file; unfortunately, due to GitHub limitations, that Notebook would have been too large to upload due to the number of generated ans printed image and caption pairs.
+While the initial plan was to contain all the information in one file; unfortunately, due to GitHub limitations, that Notebook would have been too large to upload due to the number of generated and printed image and caption pairs.
 Therefore, all training data sized but 2k and 5k were tested in separate notebooks using the same methods, adapted from \cite{ilinykh}, stored in a separate file, and imported into the notebook.
@@
 \subsection{Flickr8k (DK)}
+
+The model was trained on the Flickr8k dataset \cite{flickr8k}.
+It consists of 8,000 images that were hand-selected to depict actions and events of people or animals rather than scenery and mood.
+These images each are paired with five captions describing the image generally.
+The captions are not related to each other and were created by different annotators.
+
 \subsection{imageCLEF (MS)}
@@
 
-\subsection{Caption Evaluation (MS)}
+During the training, we tried different hyperparameters for loading the data.
+Changing the method for filtering captions that contain relations didn't provide big differences in first tests.
+Therefore, we utilized solely the simpler POS-based approach for all experiments.
+
+Concatenating all captions of the imageCLEF dataset was a too big challenge for the model and yielded worse results in first test than only using the first caption for each image.
+We concluded that the difference in syntax (more complex and longer sentences) as well as in semantics (more detailed descriptions) were too big to the Flickr8k dataset, the model was pre-trained on, but this hypothesis would need to be tested further in the future.
+For our experiments, we only used the first caption.
+
+For the experiments, the variable parameters were the \emph{number of samples} as well as the \emph{unknown filter}.
+The model was fine-tuned on a sample size (training set size) of 100 (80), 200 (160), 500 (400), 1,000 (800), 2,000 (1,600) and 5,000 (4,000).
+For each of these sample sizes, four different unknown filter thresholds were used: 0.1, 0.15, 0.2 and 1.
 
@@
 This process continued in the "500" and "1000" models with fewer and fewer captions being generated successfully, and those generated ones being absolutely nonsensical.
-Finally in the "2k" and "5k" models most captions could not be generated.\footnote{MS} \textit{This part can be moved somewhere else if it fits better for your hyperparameters section.}
+Finally, in the "2k" and "5k" models most captions could not be generated.\footnote{MS}
+
+While the human evaluation is the focus of this experiment, it is still insightful to also check, how the model fares during training and validation, more specifically, how the BLEU score changes.
+Interestingly, the models with few data samples achieve very low BLEU scores, while the models with more samples achieve higher BLEU scores.
+Table \ref{BLEU-sample-sizes} shows the best achieved BLEU scores during validation with an unknown filter of 0.1.
+For other filter values, the results look similar.
+This is opposite to the observation mentioned before.
+A reason for this may be the calculation of the BLEU score.
+The BLEU score only evaluates the appearance of words, in this experiment 4-grams, but does not focus explicitly on grammaticality.
+A lower BLEU score, but grammatical correct sentences may mean that the model doesn't learn to relate the new captions to the image features, but that it still relies on the pre-trained generations.
+On the other hand, a higher BLEU score with low grammaticality may hint to the fact, that the model is now able to generate the correct words, but looses the pre-trained ability to understand syntax.
+
+\begin{table}[h]
+    \begin{center}
+        \begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|}
+            \hline \bf number of samples & \bf highest BLEU score & \bf fine-tuned epoch \\ \hline
+            100                          & 0.065                  & 13                   \\
+            200                          & 0.091                  & 9                    \\
+            500                          & 0.146                  & 19                   \\
+            1000                         & 0.165                  & 18                   \\
+            2000                         & 0.201                  & 13                   \\
+            5000                         & 0.222                  & 9                    \\
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{\label{BLEU-sample-sizes} BLEU scores for different sample sizes (unknown filter 0.1) }
+\end{table}
+
+Looking at the unknown filter, higher thresholds produce in general better BLEU scores.
+Table \ref{BLEU-unknown-filter} shows the highest BLEU score for different unknown filter with 100 samples.
+This applies especially to the models with a larger number of samples, but the effect can be seen in all models.
+A reason for this may be the higher number of unknown tokens in the training data.
+Since, the model is also trained on generating unknown tokens, a higher share of unknown tokens in the training data will result in a higher share in the generated caption.
+Both sentences are therefore getting more similar, which results in a higher BLEU score.
+A solution for this problem could be excluding the unknown token from the model's vocabulary.
+
+\begin{table}[h]
+    \begin{center}
+        \begin{tabular}{|p{1.9cm}|p{2cm}|p{2cm}|}
+            \hline \bf unknown filter & \bf highest BLEU score & \bf fine-tuned epoch \\ \hline
+            0.1                       & 0.065                  & 13                   \\
+            0.15                      & 0.104                  & 18                   \\
+            0.2                       & 0.176                  & 12                   \\
+            1                         & 0.162                  & 19                   \\
+            \hline
+        \end{tabular}
+    \end{center}
+    \caption{\label{BLEU-unknown-filter} BLEU scores for different unknown filters (number of samples: 100) }
+\end{table}
+
+\subsection{Caption Evaluation (MS)}
 
@@
 As can be noticed in Table 1, the majority of the "good" captions for the unfiltered imageCLEF set were generated by the original, non-fine-tuned model, or by the one with a 10\% filter.
-Some captions from the model fine-tuned on nonfiltered data turned out to have been rated high as well.
+Some captions from the model fine-tuned on non-filtered data turned out to have been rated high as well.
 Finally, other models' captions were rated as good when they overlapped with the original model's caption.
@@
             2448210587[...]  & original (33.3\%)                  \\
-2445654384[...] & 100\_0.1 (50\%) or 100\_0.2 (50\%)\\
+            2445654384[...]  & 100\_0.1 (50\%) or 100\_0.2 (50\%) \\
             \hline
@@
 In addition, half of the respondents pointed out that the captions tend to include people in the descriptions even when there are no people in the images.
-It seems that the impressions of the captions in the last content section were better: while people said that they were still "strange," "weird," or "confusing," they also admitted that the verbs in the sentences tended to be accurate descriptions of the events in the images, and that they were overall "somewhat accurate." The person who expressed their feelings of anxiety previously now said that they "started to get used to them," which could also be a consequence of there being a smaller mis-match between the images and the captions.
+It seems that the impressions of the captions in the last content section were better: while people said that they were still "strange," "weird," or "confusing," they also admitted that the verbs in the sentences tended to be accurate descriptions of the events in the images, and that they were overall "somewhat accurate." The person who expressed their feelings of anxiety previously now said that they "started to get used to them," which could also be a consequence of there being a smaller mismatch between the images and the captions.
 
@@
 
-\textit{Here we compare our results to our initial expectations and decide how they answered questions, we also contrast them with prior work (if there is any).
-Highlight why what we did was relevant to the field.}
+As seen in this project, domain adaption for a pretrained image captioning model can be very challenging.
+It was surprising, to which extent the pre-trained model was disrupted when fine-tuned on data from another domain.
+Even though the dataset for fine-tuning was very different than the dataset for pre-training, we hoped that the fine-tuned model would still perform better on the new domain than the original.
+This was not the case for multiple reasons, mentioned above.
+At least, we expected it to be able to generate sensible captions for most of the images and that it would only produce big mistakes for a smaller part.
+With the rather opposite results discussed before, two major challenges for domain adaption got highlighted.
+First, the model can't adapt well to a different structure of captions.
+This results in ungrammatical sentences, when fine-tuned on many samples from the new domain.
+Secondly, the model is not able to adapt to a different topic of images.
+That results in 'reusing' old captions on new images and not learning its image features, when fine-tuned on few samples.
+For a deeper analysis, it may be helpful to first fine-tune the model on a domain that is closer in one of these aspects, for example a domain, that uses a similar way of describing the images, but focuses on topics.
+This could test the hypothesis made in this paper.
 
@@
 We have determined that if the domains are too different and too much fine-tuning data is fed to the model, the generation will start to completely fail or focus only on a very narrow range of words, producing very ungrammatical captions.
-We have also collected human judgements and determined that while fine-tuning does slighly increase the quality of the captions on the new dataset, all of them, regardless of the model, were still considered weird, mismatched, or unnatural, and that the structure of the sentences in the original dataset may have a large influence on how they are generated even after fine-tuning (e.g.
+We have also collected human judgements and determined that while fine-tuning does slightly increase the quality of the captions on the new dataset, all of them, regardless of the model, were still considered weird, mismatched, or unnatural, and that the structure of the sentences in the original dataset may have a large influence on how they are generated even after fine-tuning (e.g.
 trying to feature people in every caption).
@@
 Finally, it would be interesting to explore other potential ways of dealing with the vocabulary mismatch between different datasets.
-
-\begin{table}
-\centering
-\begin{tabular}{lc}
-\hline
-\textbf{This is} & \textbf{a table}\\
-\hline
-a table & with things \\
-and even & more things  \\\hline
-\end{tabular}
-\caption{Example table.}
-\label{tab:accents}
-\end{table}
 

### 2023-01-09 code/dataset.py

@@ -47,7 +47,6 @@ class Sample:
         encoded_caption (torch.LongTensor): The caption encoded using some wordmap.
         image (torch.FloatTensor): A tensor representation of the image associated with the sample.
     '''
-    # by Dominik
     image_id: str
     caption: str
     image_path: str

### 2023-01-09 code/train.py

@@ -27,7 +27,7 @@ training_parameters = {
 }
 
 class CaptionTrainer:
-    '''Most of the code by Nikolai Ilinykh, adapted into a class by Dominik.  TODO: make sure that is correct!!!
+    '''Most of the code by Nikolai Ilinykh, adapted into a class by Dominik.
     This class allows for defining how a model should be trained and then providing methods for carrying that training out.
 
     Attributes:

### 2023-01-09 paper/paper.tex
Final proofreading of the paper

@@
 We discover that differences in vocabulary play a major role and that providing too much fine-tuning data can cause the whole system to fail.
-We conclude that domain adaptation can be difficult to carry out, but perhaps worth attempting in certain cases.
+We conclude that domain adaptation can be difficult to carry out, but perhaps worth attempting in certain cases (MS).
 
@@
 In Section 4 we discuss how they answer our questions and how they relate to previous work.
-Finally, in Section 5 we offer conclusions and suggestions for future projects on this topic.\footnote{(MS)}
+Finally, in Section 5 we offer conclusions and suggestions for future projects on this topic.
 
@@
 Within this section we will describe the resources that we used in the project.
-We will specify the characteristics of the image captioning model, the two datasets that were used to train or fine-tune it, and the questionnaire we utilized to elicit human judgements of the best captions.\footnote{(MS)}
+We will specify the characteristics of the image captioning model, the two datasets that were used to train or fine-tune it, and the questionnaire we utilized to elicit human judgements of the best captions (MS)
 
@@
 
-The goal of this project was to analyze how the generation of captions changes, when an already pre-trained model is fine-tuned on new data from a different domain.
+The goal of this project was to analyze how the generation of captions changes when an already pre-trained model is fine-tuned on new data from a different domain.
 For this, we utilized code written by \cite{ilinykh} that implemented a model described by \cite{xu2015attend}.
@@
 
-\textit{Write about the division into files, a bit what's what, what we added and what was there.
-I did the evaluation part.
-I doubt we need to describe it in much detail, can always say that the readme has instructions and all the elements are documented, but I'd still have a bit of a summary of what we did here.}
-
 Loading the data was done in a \emph{PyTorch Dataset} class.
@@
 Here, we implemented two different methods.
-One is filtering the captions based on the POS and the other is a rule-based approach taken from \cite{Ghanimifard2018}.
+One is filtering the captions based on the POS tag, and the other is a rule-based approach taken from \cite{Ghanimifard2018}.
 Furthermore, a total number of images that should be loaded can be defined.
@@
 Since the imageCLEF dataset only consists of one independent caption instead of five as the Flickr8k dataset, the BLEU score is calculated only with one reference caption.
-The model is trained as long as the BLEU score doesn't improve for 20 epochs; the model with the best BLEU score is saved and used for testing and evaluating.
+The model is trained as long as the BLEU score doesn't improve for 20 epochs; the model with the best BLEU score is saved and used for testing and evaluating (DK).
 
@@
 Other than relatively minor fixes that were required to make the code run, we decided to manually exclude the UNK token from the generation process by enforcing the lowest possible probability on it in the beam search.
-While this meant that we could not evaluate to what extent the models favored generating that token, it did allow us to obtain captions that sounded more natural, which was paramount for the subsequent human evaluation.\footnote{(MS)}
+While this meant that we could not evaluate to what extent the models favored generating that token, it did allow us to obtain captions that sounded more natural, which was paramount for the subsequent human evaluation (MS).
 
@@
 
-\textit{Here we should summarize our results, in terms of:
-    \begin{enumerate}
-        \item where did the performance exceed the best prior performance (we got BEST checkpoints?) \textbf{DK}
-        \item which models actually generated reasonable captions? where did they start falling apart? \textbf{MS or DK}
-    \end{enumerate}
-    Include figures and tables as much as we can/as much as is reasonable.}
-
 \subsection{Hyperparameters (DK)}
@@
 Changing the method for filtering captions that contain relations didn't provide big differences in first tests.
-Therefore, we utilized solely the simpler POS-based approach for all experiments.
+Therefore, we utilized solely the simpler POS-based approach that we developed ourselves for all experiments.
 
-Concatenating all captions of the imageCLEF dataset was a too big challenge for the model and yielded worse results in first test than only using the first caption for each image.
+Concatenating all captions of the imageCLEF dataset was a too big of a challenge for the model and yielded worse results in first test than only using the first caption for each image.
 We concluded that the difference in syntax (more complex and longer sentences) as well as in semantics (more detailed descriptions) were too big to the Flickr8k dataset, the model was pre-trained on, but this hypothesis would need to be tested further in the future.
-For our experiments, we only used the first caption.
+For our experiments, we only used the first section of the caption.
 
@@
 The model was fine-tuned on a sample size (training set size) of 100 (80), 200 (160), 500 (400), 1,000 (800), 2,000 (1,600) and 5,000 (4,000).
-For each of these sample sizes, four different unknown filter thresholds were used: 0.1, 0.15, 0.2 and 1.
+For each of these sample sizes, four different unknown filter thresholds were used: 0.1, 0.15, 0.2 and 1 (DK).
 
-As described previously in the Materials and Methods section, the only models that consistently produced captions for almost all test images were the original and the "100" models.
+As mentioned previously in the Materials and Methods section, the only models that consistently produced captions for almost all test images were the original and the "100" models.
 The "200" models still generated most of the captions, but many of them started being rather nonsensical or non-grammatical.
 This process continued in the "500" and "1000" models with fewer and fewer captions being generated successfully, and those generated ones being absolutely nonsensical.
-Finally, in the "2k" and "5k" models most captions could not be generated.\footnote{MS}
+Finally, in the "2k" and "5k" models most captions could not be generated (MS).
 
@@
 The BLEU score only evaluates the appearance of words, in this experiment 4-grams, but does not focus explicitly on grammaticality.
-A lower BLEU score, but grammatical correct sentences may mean that the model doesn't learn to relate the new captions to the image features, but that it still relies on the pre-trained generations.
+A lower BLEU score, but grammatical correct sentences may mean that the model does not learn to relate the new captions to the image features, but that it still relies on the pre-trained generations.
 On the other hand, a higher BLEU score with low grammaticality may hint to the fact, that the model is now able to generate the correct words, but looses the pre-trained ability to understand syntax.
@@
 Both sentences are therefore getting more similar, which results in a higher BLEU score.
-A solution for this problem could be excluding the unknown token from the model's vocabulary.
+A solution for this problem could be excluding the unknown token from the model's vocabulary (DK).
 
@@
 Therefore, the results presented here are divided between the sets and include only the information on which model's caption was deemed the best, as well as how many people voted for it.
-Table 1 contains the information for the unfiltered imageCLEF set, Table 2 - for the filtered one, and Table 3 for Flickr8k.
+Table \ref{unfiltered-table} contains the information for the unfiltered imageCLEF set, Table \ref{filtered-table} - for the filtered one, and Table \ref{flickr-table} for Flickr8k.
 Image names are the file names without the .jpg extension.
@@
 
-As can be noticed in Table 1, the majority of the "good" captions for the unfiltered imageCLEF set were generated by the original, non-fine-tuned model, or by the one with a 10\% filter.
+As can be noticed in Table \ref{unfiltered-table}, the majority of the "good" captions for the unfiltered imageCLEF set were generated by the original, non-fine-tuned model, or by the one with a 10\% filter.
 Some captions from the model fine-tuned on non-filtered data turned out to have been rated high as well.
@@
 
-When it comes to the filtered imageCLEF test set (Table 2), where a maximum of 10\% tokens in the original caption could be unknown tokens, the spectrum of what models' captions were considered good is even wider.
+When it comes to the filtered imageCLEF test set (Table \ref{filtered-table}), where a maximum of 10\% tokens in the original caption could be unknown tokens, the spectrum of what models' captions were considered good is even wider.
 There was only one case where one caption was voted best by all the participants.
@@
 
-Finally, when it comes to the original Flickr8k-based test set, as seen in Table 3, the original model without fine-tuning performed best, although there were instances where the 10\% or 20\% fine-tuned models generated better captions.
+Finally, as for the original Flickr8k-based test set, as seen in Table \ref{flickr-table}, the original model without fine-tuning performed best, although there were instances where the 10\% or 20\% fine-tuned models generated better captions.
 Many of the original model's captions were also chosen unanimously.
@@
 As seen in this project, domain adaption for a pretrained image captioning model can be very challenging.
-It was surprising, to which extent the pre-trained model was disrupted when fine-tuned on data from another domain.
+It was surprising to what extent the pre-trained model was disrupted when fine-tuned on data from another domain.
 Even though the dataset for fine-tuning was very different than the dataset for pre-training, we hoped that the fine-tuned model would still perform better on the new domain than the original.
 This was not the case for multiple reasons, mentioned above.
-At least, we expected it to be able to generate sensible captions for most of the images and that it would only produce big mistakes for a smaller part.
+At the very least, we expected it to be able to generate sensible captions for most of the images and that it would only produce big mistakes for a smaller part.
 With the rather opposite results discussed before, two major challenges for domain adaption got highlighted.
-First, the model can't adapt well to a different structure of captions.
+First, the model cannot adapt well to a different structure of captions.
 This results in ungrammatical sentences, when fine-tuned on many samples from the new domain.
@@
 Throughout this project we have explored and tested how a CNN and LSTM-based image caption generation model could be fine-tuned, or adapted to a new domain, as represented by a somewhat different image-caption dataset.
-We have identified points that can be especially problematic, such as differences in the vocabulary and the subsequent generation of UNK tokens in the captions.
-We have determined that if the domains are too different and too much fine-tuning data is fed to the model, the generation will start to completely fail or focus only on a very narrow range of words, producing very ungrammatical captions.
+We have identified points that can be especially problematic, such as differences in the vocabulary and sentence structure, and the subsequent generation of UNK tokens in the captions as well as ungrammatical sentences.
+We have determined that if the domains are too different and too much fine-tuning data is fed to the model, the generation will start to completely fail or focus only on a very narrow range of words.
 We have also collected human judgements and determined that while fine-tuning does slightly increase the quality of the captions on the new dataset, all of them, regardless of the model, were still considered weird, mismatched, or unnatural, and that the structure of the sentences in the original dataset may have a large influence on how they are generated even after fine-tuning (e.g.
@@
 It would also be fascinating to see similar research pertaining to other model architectures, especially trying to identify the optimal hyperparameters for fine-tuning transformer-based image caption generation models.
+Of course, following up on the idea from the Discussion section, to verify our conclusions from the results by fine-tuning on a more similar dataset is another topic worth investigating.
 Finally, it would be interesting to explore other potential ways of dealing with the vocabulary mismatch between different datasets.
@@
 
-\textit{Here we have our statement of who did what.}
 While really detailed information on who developed what is available in the code, the code is not the only thing that we shared work on.
@@
     \item \textbf{Writing up:} both Dominik and Maria.
-Abstract by Maria (cannot put the initials there).
-While the initials signify the person who wrote most of the section, small changes by the other person are possible.
-Some sections' authorship is shared; in that case a footnote is put at the end of the part written by one person.
+While the initials signify the person who wrote most of the section, small changes by the other person are possible - and are marked by (initials) at the end of a paragraph.
+This is done especially if a certain paragraph was moved from one section to the other.
+Otherwise, for sections where authorship is more evenly shared, the same method of marking contribution is employed.
     \item \textbf{Repository management:} Maria with Dominik's help.
+    \item \textbf{Presentation:} Maria with Dominik's help.
 \end{enumerate}
